---
title: "ds202_lab5"
author: "Tyren Teav"
date: "4/15/2021"
output: html_document
---
Problem 1:

Saturdays have the most accidents than any other day of the week, followed by Sunday which makes sense. Weekends the most things are going on, people are going out the most, likely the most activity. 
```{r}
library(ggplot2)
library(tidyverse)

acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)

per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

acc %>%
  group_by(acc$DAY_WEEK) %>%
  summarize(
    count = length(DAY_WEEK)
  ) %>%
  arrange(count)




```


Problem 2

```{r}

injury <- subset(per, INJ_SEV >= 1)
head(injury)



```

Problem 3
I used INJ_SEV like problem 2 to represent the fatalities 
```{r}
library(dplyr)


per2 <- per %>%
  group_by(STATE, MAKE) %>%
  summarize(
    deaths = sum(INJ_SEV)
  ) %>%
  arrange(STATE)

per3 <- na.omit(per2)

per4 <- per3 %>%
  group_by(STATE) %>%
  top_n(1,deaths)

per4
```

Problem 4

The most dangerous vehicle is described as the vehicle make that has caused the most INJ_SEV from problem 2 and 3. The map shows that there are two main makes that have cause the most deaths, being ,make 20 and 12.
```{r}

library(readxl)
library(tidyverse)
states <- map_data('state')
US <- readxl::read_xlsx("/Users/Tyren/Desktop/DS_202/United_States.xlsx")

ggplot(states, aes(x = long, y = lat)) + geom_polygon(aes(group=group))

US$`State Name` <- tolower(US$`State Name`)
text <- full_join(US, states, by = c("State Name" = "region"))

text1 <- select(text, "State Name", "State Code", "long", "lat")

per4$STATE = as.character(per4$STATE)

locations <- text1 %>%
  group_by(`State Name`, `State Code`) %>%
  summarize(
    long = mean(long),
    lat = mean(lat)
  )

text2 <- full_join(locations, per4, by = c('State Code' = "STATE"))

ggplot(states, aes(x=long, y = lat)) + geom_polygon(aes(group=group)) + geom_text(aes(label= MAKE), data = text2, color = "Green")






```